SU($n$) spin models near the Heisenberg critical point

### 2020-03-24 collective_sun/sun_perturbation.tex
major changes allowing for the consideration of more general multi-body operators

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\diag}{diag}
\newcommand{\ul}{\underline}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator of the general form \red{[TODO: consider more
  general sums of such operators]}
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all such choices of $M$ elements from a set of size $N$; and
$\m w$ is a dimension-$M$ tensor with scalar entries $w_k$.  Even
though only ${N\choose M}$ entries of $\m w$ are used to define
$X_{\m w}$ in \eqref{eq:pert}, for later convenience we define $w_k$
for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric
with respect to any permutation of its indices, and that $w_k=0$ for
any $k$ with repeated values.  In contrast, $X_k$ generally need not
be symmetric with respect to permutations of $k$, although we will
still define $X_k=0$ for any $k$ with repeated values.  A chiral spin
model, for example, may have two-body interactions of the form
$X_{pq}=s_\x^{(p)}s_\y^{(q)}-s_\y^{(p)}s_\x^{(q)}$, which are not
invariant under exchange of spins $p$ and $q$.

Single-body ($M=1$) operators correspond to classical external fields,
e.g.~$s_\z$ for a magnetic field, and two-body ($M=2$) operators
correspond to two-body interactions that deviate from $H_0$, such as
$s_\z\otimes s_\z$ interactions in an XXZ model\red{[TODO:cite]} near
the Heisenberg critical point.  While we would be hard-pressed to find
a physical implementation of multi-body spin interactions with $M>2$,
considering this case will be necessary to construct low-lying
eigenstates of $H_0$.  Constructing low-lying eigenstates of $H_0$
will in turn enable simulating the dynamics induced by
non-perturbative additions $X_{\m w}$ to the SU($n$) Heisenberg model,
with norms $\norm{X_{\m w}}\sim\Delta_{\t{gap}}$.

When $X_{\m w}$ is sufficiently small, we can treat its effect on the
PS manifold $\M_0$ perturbatively.  The effective Hamiltonian
$H_{\t{eff}}$ induced by $X_{\m w}$ on the ground-state manifold
$\M_0$ through second order in perturbation theory
is\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
  &&
  H_{\t{eff}}^{(1)} = \P_0 X_{\m w} \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w} \P_\Delta X_{\m w} \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
simplify using the permutational symmetry of $\M_0$:
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \bar w\,\col{X},
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
is an average of $w_k$ over spin choices $k\in\C_N\p{M}$; and
$\col{X}$ is the collective version of $X$, i.e.
\begin{align}
  \bar w \equiv { N \choose M }^{-1} \sum_{k\in\C_N\p{M}} w_k,
  &&
  \col{X} \equiv \sum_{k\in\C_N\p{M}} X_k,
\end{align}
An immediate consequence of \eqref{eq:H_eff_1} is the fact that
perturbing SU(2)-symmetric interactions by two-body $s_\z\otimes s_\z$
interactions yields the one-axis twisting (OAT)
Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
$H_{\t{OAT}}=\chi S_\z^2$ at first order in perturbation theory, with
the OAT strength $\chi$ equal to the mean coefficient of the two-body
$s_\z\otimes s_\z$ terms.

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  This projector essentially
picks off the part of $X_{\m w}$ that is strictly off-diagonal with
respect to the ground- and excited-state manifolds $\E_0$ and
$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
states of definite excitation energy when acting on PS states $\M_0$,
we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
  \label{eq:multi_body_eig_op}
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
symmetries of $H_0$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

Defining $\m w_\Delta$ to be the projection of the tensor $\m w$ onto
the $\Delta$-eigenspace of $\hat{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
  w_\Delta} \P_0$, and so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  &&
  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
\end{align}
which in the SU(2) case implies that an inhomogeneous magnetic field
($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $X_{\m w}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul{\M}_M$, i.e.~$\ul{\M}_M\equiv\bigoplus_{r\le M}\M_r$.  The
manifold $\ul{\M}_M$ consists of all states accessible by the action
of $r$-body operators with $r\le M$ on PS states in $\M_0$.

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert_copy}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all such choices of $M$ elements from a set of size $N$; and
$\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor with scalar entries
$w_k$.  Even though only ${N\choose M}$ entries of $\m w$ are used to
define $X_{\m w}$ in \eqref{eq:pert_copy}, for later convenience we
define $w_k$ for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is
symmetric with respect to any permutation of its indices, and that
$w_k=0$ for any $k$ with repeated values.  In contrast, $X_k$
generally need not be symmetric with respect to permutations of $k$,
although we will still define $X_k=0$ for any $k$ with repeated
values.

@@

\red{[TODO: verify that the derivation below properly accounts for
  operators $X$ that are not permutationally symmetric]}

@@

$p,q\notin k$.  When $p,q\in k$, this permutation operator swaps the
positions of $p$ and $q$ in $X_k$, but we can simply relabel
$\p{p,q}\leftrightarrow\p{q,p}$ with no consequence for $w_k$, so

@@

      p,q\notin k~\t{or}~p,q\in k}}

@@

  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\notin k~\t{or}~p,q\in k}}

@@

For fixed $k\in\C_N\p{M}$, we can simplify

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric manifold (PS)
$\M_0$.  For $p$ multi-body operators, such a product takes the form
\begin{align}
  \m X_{\m w}
  \equiv \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\C_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
$M_\alpha$-spin operators $X_\alpha$;
$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
tensors $w_\alpha$;
$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
is an ordered choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$
acts on; $\C_N\p{M_\alpha}$ is the set of all such choices of
$M_\alpha$ elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$
is a scalar entry of $w_\alpha$.  Although only ${N \choose M_\alpha}$
entries of $w_\alpha$ are used in \eqref{eq:sym_prod_start}, we define
$w_\alpha\p{k_\alpha}$ for all $k_\alpha\in\ZZ_N^{M_\alpha}$ by
enforcing that $w_\alpha\p{k_\alpha}$ is symmetric with respect to
permutations of $k_\alpha$, and that $w_\alpha\p{k_\alpha}=0$ if
$k_\alpha$ has any repeated values.  In contrast,
$X_\alpha\p{k_\alpha}$ generally need not be symmetric with respect to
permutations of $k_\alpha$, although we still enforce that
$X_\alpha\p{k_\alpha}=0$ if $k_\alpha$ has any repeated values.

>>>>

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\diag}{diag}
\newcommand{\ul}{\underline}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator $\hat\O$ of the general form
\begin{align}
  \hat\O \equiv \sum_{\p{X,\m w}\in\O} X_{\m w},
  &&
  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $\hat\O$ is constructed from the set $\O\equiv\set{\p{X,\m w}}$
of $M$-spin operators $X$ together with dimension-$M$ (i.e.~$M$-index)
tensors $\m w$; $k=\p{k_1,k_2,\cdots,k_M}$ is a list of of $M$ spins
that $X_k$ acts on; $w_k$ is a scalar entry of $\m w$; and
\begin{align}
  \D_N\p{M} \equiv
  \set{ k \in \ZZ_N^M:~\t{all elements of}~k~\t{are distinct} },
  \label{eq:off_diags}
\end{align}
is the set of all ordered choices of $M$ spins out of $N$.  The
normalization factor of $1/M!$ in \eqref{eq:pert} is included to
account for the fact that there are $M!$ terms addressing any given
choice of $M$ of spins.  Due to the permutational symmetry of the
ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$, we
will find it convenient to decompose
$\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where $\pi\p{M}$ is the
permutation group of $\ZZ_M$, and $\C_N\p{M}$ is the set of all
choices of $M$ spins out of $N$.  We can then expand
\begin{align}
  X_{\m w} = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}}.
\end{align}
For simplicity, we will assume throughout this work that $\O$ is
homogeneous in the sense that $M$ is the same for all
$\p{X,\m w}\in\O$, with the understanding that a generalization to
different $M$ for each $\p{X,\m w}\in\O$ is straightforward.

Single-body ($M=1$) operators correspond to classical external fields,
e.g.~$s_\z$ for a magnetic field, and two-body ($M=2$) operators
correspond to two-body interactions that deviate from $H_0$, such as
$s_\z\otimes s_\z$ interactions in an XXZ model\red{[TODO:cite]} near
the Heisenberg critical point.  While we would be hard-pressed to find
a physical implementation of multi-body spin interactions with $M>2$,
considering this case will be necessary to construct low-lying
eigenstates of $H_0$, which will in turn enable simulating the
dynamics induced by non-perturbative additions $\hat\O$ to the SU($n$)
Heisenberg model, with operator norms
$\norm*{\hat\O}\sim\Delta_{\t{gap}}$.

When the operator norm of $\hat\O$ is sufficiently small, namely
$\norm*{\hat\O}<\Delta_{\t{gap}}/2$, we can treat its effect on the PS
manifold $\M_0$ perturbatively.  The effective Hamiltonian
$H_{\t{eff}}$ induced by $\hat\O$ on the ground-state manifold $\M_0$
through second order in perturbation theory
is\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
  &&
  H_{\t{eff}}^{(1)} = \P_0 \hat\O \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \hat\O \P_\Delta \hat\O \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
simplify using the permutational symmetry of $\M_0$:
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \sum_{\p{X,\m w}\in\O} \bar w\,\col{X},
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
is an average of $w_k$ over all $k\in\D_N\p{M}$; and $\col{X}$ is a
permutation-symmetrized collective version of $X$, i.e.
\begin{align}
  \bar w \equiv \f1{\abs{\D_N\p{M}}} \sum_{k\in\D_N\p{M}} w_k,
  &&
  \col{X} \equiv \f1{\abs{\pi\p{M}}} \sum_{\sigma\in\pi\p{M}}
  \sum_{k\in\C_N\p{M}} X_{\sigma\p{k}},
\end{align}
where, to be explicit,
\begin{align}
  \abs{\D_N\p{M}} = N!_M,
  &&
  \pi\p{M} = M!,
\end{align}
with $N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$
denoting a falling factorial.  An immediate consequence of
\eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
interactions by two-body $s_\z\otimes s_\z$ interactions yields the
one-axis twisting (OAT) Hamiltonian\cite{kitagawa1993squeezed,
  ma2011quantum} $H_{\t{OAT}}=\chi S_\z^2$ at first order in
perturbation theory, with the OAT strength $\chi$ equal to the mean
coefficient of the two-body $s_\z\otimes s_\z$ terms.

\red{[TODO: fix remainder of this section to account for perturbations
  that do not obey permutational symmetry]}

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  For each term $X_{\m w}$ of
$\hat\O$, this projector essentially picks off the part of $X_{\m w}$
that is strictly off-diagonal with respect to the ground- and
excited-state manifolds $\E_0$ and $\E_\Delta$.  To decompose
$X_{\m w}$ into components that generate states of definite excitation
energy when acting on PS states $\M_0$, we pick an arbitrary state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
  \label{eq:multi_body_eig_op}
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
symmetries of $H_0$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

Defining $\m w_\Delta$ to be the projection of the tensor $\m w$ onto
the $\Delta$-eigenspace of $\hat{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 =
X_{\m{w}_\Delta} \P_0$, so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
  \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
  \P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
  \f{\cov\p{\m v_\Delta,\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_eff_2_1}
\end{align}
where
\begin{align}
  \cov\p{\m v,\m w} \equiv
  \f1N \sum_{p\in\ZZ_N} \p{v_p-\bar v}\p{w_p-\bar w}
\end{align}
is the covariance between $\m v$ and $\m w$.  In the SU(2) case,
\eqref{eq:H_eff_2_1} implies that an inhomogeneous magnetic field
($X\propto s_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul{\M}_M$, i.e.~$\ul{\M}_M\equiv\bigoplus_{r\le M}\M_r$.  The
manifold $\ul{\M}_M$ consists of all states accessible by the action
of $r$-body operators with $r\le M$ on PS states in $\M_0$.

@@

\red{[TODO: fix section to consider the case without permutational
  symmetry of $X_{\m w}$]}

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:pert_copy}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a list of $M$ spins that $X_k$ acts on; $\D_N\p{M}$ is defined in
\eqref{eq:off_diags}; and $\m w$ is a dimension-$M$ (i.e.~$M$-index)
tensor with scalar entries $w_k$.

@@

\red{[TODO: fix section to consider the case without permutational
  symmetry of $X_{\m w}$]}

@@

$p,q\in k$ or $p,q\notin k$, so

@@

      p,q\in k~\t{or}~p,q\notin k}}

@@

  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}}

@@

For any fixed $k\in\C_N\p{M}$, we can simplify

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric (PS) manifold
$\M_0$.  For $p$ multi-body operators, such a product takes the form
\begin{align}
  \m X_{\m w}
  \equiv \prod_{\alpha\in\ZZ_p} \f1{M_\alpha!}
  \sum_{k_\alpha\in\D_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
$M_\alpha$-spin operators $X_\alpha$;
$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
tensors $w_\alpha$;
$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
$\D_N\p{M_\alpha}$ is the set of all ordered choices of $M_\alpha$
elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$ is a
scalar entry of $w_\alpha$.  For convenience, we define
$w_\alpha\p{k_\alpha}$ and $X_\alpha\p{k_\alpha}$ for all
$k_\alpha\in\ZZ_N^{M_\alpha}$ by enforcing that
$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for all $k_\alpha$ with
repeated values.  These definitions allow us to replace sums over
$\D_N\p{M_\alpha}$ by sums over $\ZZ_N^{M_\alpha}$, thereby making the
sums over the individual indices in $k_\alpha$ independent.

To simplify the product in \eqref{eq:sym_prod_start}, we first replace
all $\D_N\p{M_\alpha}$ by $\ZZ_N^{M_\alpha}$ and rewrite the product
of sums as a sum of products:
\begin{align}
  \m X_{\m w}
  &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_1}} \cdots
  \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
  \prod_{\alpha\in\ZZ_p} w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha} \\
  &= \f1{\m M!} \sum_{k\in\ZZ_N^{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sum_of_products}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
$M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for
shorthand; and
$\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
that each $k\in\ZZ_N^{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\ZZ_N^{M_\alpha}$.
